On resolution sensitivity in the Community Atmosphere Model

### 2019-09-15 resolution.tex
edits

The vertical velocities are found to fit a power law scaling $\Delta x^n$ with $\Delta x$ the grid spacing and $n=-1$. The $n=-1$ exponent is derived from a scale analysis of the Poisson equation, indicating that pressure gradients scale like the inverse of the horizontal scale of buoyancy perturbations, $D^{-1}$, driving convergence into the air column and the vertical velocities also scale like $D^{-1}$. It has been previously shown that $D$ is set by stratiform cloud formation, which collapse to the smallest resolvable feature in the model \citep{HR2018JAMES}, the effective resolution \citep{S2011LNCSE}, and so $D \propto \Delta x$ giving $n=-1$. This scaling is consistent with studies describing the resolution sensitivity of large-eddy simulations, which use similar physical arguments, but adapted for non-hydrostatic scales \citep{WETAL1997MWR,PG2006JAS,JR2016QJRMS}.

The $n=-1$ power law scaling is potentially in conflict with the scaling proposed by \cite{RETAL2016CD}, $n=h-1$, with $h$ being half the exponent of the second-order structure function $\Delta x^{2h}$ of the horizontal wind. \cite{RETAL2016CD} propose that since the $h=\frac{1}{3}$ is supported by observations, it is an emergent constraint for $n=-\frac{2}{3}$. This is an intriguing proposal and the authors don't seek to dismiss its potential value. But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is only marginally resolved in the highest resolution simulation $\Delta x=27.8$ km ($ne120$), and CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}. The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of the kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem. For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is shallower than the slope near where the spectrum approaches the effective resolution, for all grid resolutions (Figure~~\ref{fig:2ke}). There is however, a progressive shoaling of the slope $-\beta$ with increasing resolution (Figure~~\ref{fig:2ke}b). This suggests a transition towards the $-\frac{5}{3}$ slope in the model, but seeing this process through would require running the model at higher resolutions than considered in this study.

\begin{figure}
\begin{center}
\noindent\includegraphics[width=20pc,angle=0]{figs/temp_kespectra.pdf}\\
\end{center}
\caption{(a) Kinetic energy spectrum and (b) compensated kinetic energy spectrum at the 200 hPa level in the simulations. The diagonal black line in (a) is a reference spectrum with a $-3$ slope. Compensated kinetic energy spectrum refers to the kinetic energy spectrum multiplied by $n^{-\frac{5}{3}}$, with spherical wavenumber $n$, so that horizontal lines indicate a spectrum with a $-\frac{5}{3}$ slope. In (b), thick ``caterpillars" are overlain on the portion of each spectrum pertaining to the $5-10 \Delta x$ effective resolution of the model.}
\label{fig:2ke}
\end{figure}

An increase in resolved moisture fluxes through cloud base occurs with increasing resolution, and balanced by an increase in stratiform precipitation rates. Through decomposing the stratiform precipitation into the components of the moisture flux, it is shown that the increase in stratiform precipitation rates with resolution is primarily due to the increase in resolved mass fluxes arising from the increase in resolved vertical velocities with resolution. At higher resolutions, larger stratiform precipitation rates increasingly contribute to the increase in climatological stratiform precipitation rate, with smaller magnitude precipitation rates contributing less. The increasing occurrence of larger magnitude vertical velocities and stratiform precipitation rates at higher resolution is at odds with a similar analysis by \cite{TETAL2018CD} using a similar model. The authors suggests that their use of daily averaged fields obscured the covariance between vertical velocity and moisture (i.e., moisture flux) and complicated their calculation.

@@

This study argues that the sensitivity of parameterized convection and stratiform precipitation rates to resolution are a response to the increase in magnitude of vertical velocities with resolution. This is essentially the hypothesis put forth by \cite{HR2017JCLIM} to explain resolution sensitivity, that increased vertical velocities with resolution interacts with other model components, steering the model towards a new mean state. Here, those interactions are documented explicitly. Resolution sensitivity therefore arises from the relationship between vertical velocities and resolution, a well-known scale dependency of the equations of motion \citep{O1981JAS}. This scale dependency needs to be addressed in the pursuit of scale-aware physics.

>>>>

The vertical velocities are found to fit a power law scaling $\Delta x^n$ with $\Delta x$ the grid spacing and $n=-1$. The $n=-1$ exponent is derived from a scale analysis of the Poisson equation, with pressure gradients scaling like the inverse of the horizontal scale of buoyancy perturbations, $D^{-1}$, driving convergence into the air column and vertical velocities that also scale like $D^{-1}$. It has been previously shown that $D$ is set by stratiform cloud formation, which collapse to the smallest resolvable feature in the model \citep{HR2018JAMES}, the effective resolution \citep{S2011LNCSE}, and so $D \propto \Delta x$ giving $n=-1$. This scaling is consistent with studies describing the resolution sensitivity of large-eddy simulations, which use similar physical arguments, but adapted for non-hydrostatic scales \citep{WETAL1997MWR,PG2006JAS,JR2016QJRMS}.

The $n=-1$ power law scaling is potentially in conflict with the scaling proposed by \cite{RETAL2016CD}, $n=h-1$, with $h$ being half the exponent of the second-order structure function $\Delta x^{2h}$ of the horizontal wind. \cite{RETAL2016CD} propose that since the $h=\frac{1}{3}$ is supported by observations, $n=-\frac{2}{3}$ is an emergent constraint on models. This is an intriguing proposal and the authors don't seek to dismiss its potential value. But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is only marginally resolved in the highest resolution simulation $\Delta x=27.8$ km ($ne120$), and CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}. The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of the kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem. For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is shallower than the slope near where the spectrum approaches the effective resolution, for all grid resolutions (Figure~~\ref{fig:2ke}). There is however, a progressive shoaling of the slope $-\beta$ with increasing resolution (Figure~~\ref{fig:2ke}b). This suggests a transition towards the $-\frac{5}{3}$ slope in the model, but seeing this process through would require running the model at higher resolutions than considered in this study.

\begin{figure}
\begin{center}
\noindent\includegraphics[width=20pc,angle=0]{figs/temp_kespectra.pdf}\\
\end{center}
\caption{(a) Kinetic energy spectrum and (b) compensated kinetic energy spectrum at the 200 hPa level in the simulations. The diagonal black line in (a) is a reference spectrum with a $-3$ slope. Compensated kinetic energy spectrum refers to the kinetic energy spectrum multiplied by $k^{-\frac{5}{3}}$, with spherical wavenumber $k$, so that horizontal lines indicate a spectrum with a $-\frac{5}{3}$ slope. In (b), solid circles are overlain on each spectrum pertaining to $10 \Delta x$, a conservative estimate for the effective resolution of the model.}
\label{fig:2ke}
\end{figure}

An increase in resolved moisture fluxes through cloud base occurs with increasing resolution, and balanced by an increase in stratiform precipitation rates. Through decomposing the stratiform precipitation into the components of the moisture flux, it is shown that the increase in stratiform precipitation rates with resolution is primarily due to the increase in resolved mass fluxes arising from the increase in resolved vertical velocities with resolution. At higher resolutions, larger stratiform precipitation rates increasingly contribute to the increase in climatological stratiform precipitation rate, with smaller magnitude precipitation rates contributing less. The increasing occurrence of larger magnitude vertical velocities and stratiform precipitation rates at higher resolution is at odds with a similar analysis by \cite{TETAL2018CD} using a similar model. The authors suggests that their use of daily averaged fields obscured the covariance between vertical velocity and moisture (i.e., moisture flux), complicating the analysis.

@@

This study argues that the sensitivity of parameterized convection and stratiform precipitation rates to resolution are a response to the increase in magnitude of vertical velocities with resolution. This is essentially the hypothesis put forth by \cite{HR2017JCLIM} to explain resolution sensitivity, that increased vertical velocities with resolution interacts with other model components, steering the model towards a new mean state. Here, those interactions are documented explicitly. Resolution sensitivity therefore arises from the relationship between vertical velocities and resolution, a well-known scale dependency of the equations of motion \citep{O1981JAS}.
